Implementing \Kwd{DNS} in \Kwd{C0}

### 2017-12-08 paper.tex
start revamping paper

\documentclass{article}
\usepackage[tt=false]{libertine}
\usepackage{mathpartir}
\usepackage{amsthm}
\usepackage[colorlinks=true]{hyperref}
\usepackage{bytefield}
\usepackage{fullpage}
\usepackage{listings}
\usepackage{microtype}

@@

\title{Implementing DNS in C0}
\author{Farzaneh Derakhshan \and Klaas Pruiksma \and Jonathan Sterling}

@@

\begin{abstract}
  We are implementing a \Kwd{DNS} server in the \Kwd{C0} language,
  which we intend to evaluate for both safety and performance.
\end{abstract}

@@

\subsection{C0 and Concurrent C0}\label{sec:c0}

The \Kwd{C0} programming language is a restricted subset of \Kwd{C}
that provides implicit memory safety to
programmers~\cite{pfenning:c0-reference}. Restrictions on pointer
arithmetic and casting, for example, allow the language to detect and
prevent any attempt to access an array out of its bounds. Furthermore,
there is no need for error-prone manual memory management, since
\Kwd{C0} is itself garbage-collected. Further safety conditions can be
enforced with contracts, which are a class of assertions including
loop invariants, pre- and post-conditions, and dynamically checkable
correctness conditions for programs.

\Kwd{Concurrent C0} (\Kwd{CC0}) is an extension of \Kwd{C0} that
offers concurrency based on message passing by providing processes as
the units of concurrent execution and channels as their message
buffers~\cite{willsey-prabhu-pfenning:2016}. Concurrent processes
communicate over channels using a protocol that is disciplined by
session types~\cite{caires-pfenning-toninho:2013,
  balzer-pfenning:2017}. In this protocol, each channel is typed such
that it can either send or receive a message and, thus, it
deliberately models the client/provider relation.

However, the protocol is also able to characterize processes that are
not completely sequential, and branch into different sequences as they
encounter choices. This typing discipline also guarantees that data is
sent and received in the correct order and by the correct processes
(an invariant called \emph{session fidelity}), and that the
communication is always one-to-one.

@@

@@

\section{Goals and Progress}

\subsection*{75\%}

Our 75\% goal is to develop a basic \Kwd{DNS} nameserver in \Kwd{C0}. In order to reduce the scope of this project, we restrict the types of resource data (and by extension, the types of queries) that we handle, and we will only provide support for \Kwd{UDP}-based queries. These two restrictions will simplify implementation, but leave enough functionality that we can evaluate our implementation both for correctness and for performance relative to other DNS implementations, as long as we restrict the test queries to those we have implemented.

The bulk of this goal falls into the two components described in \S\ref{sec:our-project}, along with a small ``main'' component that uses these to actually handle queries. Currently, we have some unmarshalling functions and about a third of the network library implemented. We expect that the remainder of this goal should go fairly quickly --- marshalling is no harder than unmarshalling, and the network interface is relatively straightforward to implement once we have decided on what functions are necessary and what their arguments should be.

\subsection*{100\%}

To extend the above to a 100\% goal, we plan to extend the functionality of our nameserver by adding caching of requests and potentially adding to the types of resource data we support. While a nameserver without caching is not particularly useful, we believe that the portion of code most likely to contain memory errors is in some combination of reading data from the network and parsing that data --- in particular, dealing with message compression. As such, the 75\% goal is a proof of concept that demonstrates that these portions can be implemented safely, while the 100\% goal extends that proof of concept to be more useable (though still more limited than commercially available nameservers). We expect that in going from our 75\% version to the 100\% version, performance (and the scope of allowable queries) will increase and that we will be able to leave our arguments for correctness unchanged.

\subsection*{125\%}

We have two potential 125\% goals, depending on the outcome of our 100\% goal. If our nameserver has good performance relative to its competitors, we will work to provide better justification for the correctness of our nameserver, ideally arriving at a proof that statically verifies that all of our memory accesses are safe. In doing this, we will be able to safely turn off bounds checking when running our nameserver, increasing performance through verifying correctness. If instead our performance is very poor (or not helped much by turning off bounds checking), we will work to optimize our nameserver based on the results of performance testing. It is difficult to say what specifically we will do, but we plan to take the general approach of using software like \texttt{perf} to determine what portions of our code are bottlenecks, and trying to make those paths faster.

\section{Evaluation Plan}

Our evaluation breaks down into two parts --- one on the safety, and
one on performance. As safety is a primary goal of this project, we
plan to evaluate our code for safety. A large part of this will come
for free from the memory safety guarantees of \Kwd{C0}, with the rest
coming from a careful analysis of our \Kwd{C} code. This does have the
disadvantage that \Kwd{C0}'s memory safety is checked dynamically,
which incurs performance overhead. Ideally, as described in our 125\%
goal, we would like to have a more static verification that memory
safety holds, both to provide better performance, and as it provides
better results than dynamic checking, since if a dynamic check fails,
the program halts, which is less than ideal.

For performance analysis, we plan to evaluate our nameserver for
latency and throughput, as these appear to be the most important
metrics for \Kwd{DNS} performance. We plan to do this using one of two
pieces of software\cite{github:dnsperf1, github:dnsperf2}
(confusingly, both named \texttt{dnsperf}), which read queries from a
file and then send these requests to a nameserver. The tools mentioned
in~\cite{github:dnsperf2} seem better-suited to our needs, as those
of~\cite{github:dnsperf1} appear to not give latency information in a
useful fashion.

We will compare our results (at various load levels) to existing
\Kwd{DNS} software (though we have yet to determine which specific
software to take as our baseline). One remaining difficulty in
evaluation lies in figuring out what sort of query workload to use ---
the performance analysis tools we have looked at suggest using an
input file with somewhere on the order of one million queries, which
is more than can feasibly be generated by hand. However, there does
not appear to be a standard benchmark for this testing as there are
for many other systems. We will likely generate large test files in a
randomized way, but do not have a good sense of how to make these
files accurately reflect realistic usage.

@@

\begin{lstlisting}
int parse_domain_name(int cursor, int[] data, domain_name_t* dest) {
  int* len = alloc(int);

  int orig_cursor = cursor;
  cursor = read_bits(cursor, data, 8, len);

  if (*len == 0) {
    // case: done
    *dest = domain_nil();
    return cursor;
  }

  if (*len >= 0xc0) {
    // case: pointer
    int* ptr = alloc(int);
    // read the remainder of the first 16 bits as a pointer to an offset in octets
    cursor = read_bits(orig_cursor + 2, data, 14, ptr);
    // read the last label of the domain name from the pointer:
    parse_domain_name(*ptr * 8, data, dest);
    return cursor;
  }

  // case: ordinary label
  // now '*len' specifies the number of octets in the label
  char[] chars = alloc_array(char, *len);
  for (int i = 0; i < *len; i++) {
    int* octet = alloc(int);
    cursor = read_bits(cursor, data, 8, octet);
    chars[i] = char_chr(*octet);
  }

  string lbl = string_from_chararray(chars);

  *dest = domain_cons(lbl, *dest);
  return parse_domain_name(cursor, data, dest);
}
\end{lstlisting}

>>>>

\documentclass{article}
\usepackage[tt=false]{libertine}
\usepackage{mathpartir}
\usepackage{amsthm}
\usepackage[colorlinks=true]{hyperref}
\usepackage{bytefield}
\usepackage{listings}
\usepackage{microtype}

@@

\title{Implementing \Kwd{DNS} in \Kwd{C0}}
\author{Farzaneh Derakhshan \and Klaas Pruiksma \and Jonathan Sterling}

@@

\begin{abstract}
  We have implemented a \Kwd{DNS} server in the \Kwd{C0} language, and
  evaluated it for both safety and performance.
\end{abstract}

@@

\subsection{C0}\label{sec:c0}

The \Kwd{C0} programming language is a restricted subset of \Kwd{C}
that provides implicit memory safety to
programmers~\cite{pfenning:c0-reference}. Restrictions on pointer
arithmetic and casting, for example, allow the language to detect and
prevent any attempt to access an array out of its bounds. Furthermore,
there is no need for error-prone manual memory management, since
\Kwd{C0} is itself garbage-collected.

Further safety conditions can be enforced with contracts, which are a
class of assertions including loop invariants, pre- and
post-conditions, and dynamically checkable correctness conditions for
programs.

% \Kwd{Concurrent C0} (\Kwd{CC0}) is an extension of \Kwd{C0} that
% offers concurrency based on message passing by providing processes as
% the units of concurrent execution and channels as their message
% buffers~\cite{willsey-prabhu-pfenning:2016}. Concurrent processes
% communicate over channels using a protocol that is disciplined by
% session types~\cite{caires-pfenning-toninho:2013,
%   balzer-pfenning:2017}. In this protocol, each channel is typed such
% that it can either send or receive a message and, thus, it
% deliberately models the client/provider relation.

% However, the protocol is also able to characterize processes that are
% not completely sequential, and branch into different sequences as they
% encounter choices. This typing discipline also guarantees that data is
% sent and received in the correct order and by the correct processes
% (an invariant called \emph{session fidelity}), and that the
% communication is always one-to-one.

@@

\paragraph{Nail} TODO (Jon) evaluate Nail.
\nocite{bangert:2014}

@@

% \section{Goals and Progress}

% \subsection*{75\%}

% Our 75\% goal is to develop a basic \Kwd{DNS} nameserver in \Kwd{C0}. In order to reduce the scope of this project, we restrict the types of resource data (and by extension, the types of queries) that we handle, and we will only provide support for \Kwd{UDP}-based queries. These two restrictions will simplify implementation, but leave enough functionality that we can evaluate our implementation both for correctness and for performance relative to other DNS implementations, as long as we restrict the test queries to those we have implemented.

% The bulk of this goal falls into the two components described in \S\ref{sec:our-project}, along with a small ``main'' component that uses these to actually handle queries. Currently, we have some unmarshalling functions and about a third of the network library implemented. We expect that the remainder of this goal should go fairly quickly --- marshalling is no harder than unmarshalling, and the network interface is relatively straightforward to implement once we have decided on what functions are necessary and what their arguments should be.

% \subsection*{100\%}

% To extend the above to a 100\% goal, we plan to extend the functionality of our nameserver by adding caching of requests and potentially adding to the types of resource data we support. While a nameserver without caching is not particularly useful, we believe that the portion of code most likely to contain memory errors is in some combination of reading data from the network and parsing that data --- in particular, dealing with message compression. As such, the 75\% goal is a proof of concept that demonstrates that these portions can be implemented safely, while the 100\% goal extends that proof of concept to be more useable (though still more limited than commercially available nameservers). We expect that in going from our 75\% version to the 100\% version, performance (and the scope of allowable queries) will increase and that we will be able to leave our arguments for correctness unchanged.

% \subsection*{125\%}

% We have two potential 125\% goals, depending on the outcome of our 100\% goal. If our nameserver has good performance relative to its competitors, we will work to provide better justification for the correctness of our nameserver, ideally arriving at a proof that statically verifies that all of our memory accesses are safe. In doing this, we will be able to safely turn off bounds checking when running our nameserver, increasing performance through verifying correctness. If instead our performance is very poor (or not helped much by turning off bounds checking), we will work to optimize our nameserver based on the results of performance testing. It is difficult to say what specifically we will do, but we plan to take the general approach of using software like \texttt{perf} to determine what portions of our code are bottlenecks, and trying to make those paths faster.

% \section{Evaluation Plan}

% Our evaluation breaks down into two parts --- one on the safety, and
% one on performance. As safety is a primary goal of this project, we
% plan to evaluate our code for safety. A large part of this will come
% for free from the memory safety guarantees of \Kwd{C0}, with the rest
% coming from a careful analysis of our \Kwd{C} code. This does have the
% disadvantage that \Kwd{C0}'s memory safety is checked dynamically,
% which incurs performance overhead. Ideally, as described in our 125\%
% goal, we would like to have a more static verification that memory
% safety holds, both to provide better performance, and as it provides
% better results than dynamic checking, since if a dynamic check fails,
% the program halts, which is less than ideal.

% For performance analysis, we plan to evaluate our nameserver for
% latency and throughput, as these appear to be the most important
% metrics for \Kwd{DNS} performance. We plan to do this using one of two
% pieces of software\cite{github:dnsperf1, github:dnsperf2}
% (confusingly, both named \texttt{dnsperf}), which read queries from a
% file and then send these requests to a nameserver. The tools mentioned
% in~\cite{github:dnsperf2} seem better-suited to our needs, as those
% of~\cite{github:dnsperf1} appear to not give latency information in a
% useful fashion.

% We will compare our results (at various load levels) to existing
% \Kwd{DNS} software (though we have yet to determine which specific
% software to take as our baseline). One remaining difficulty in
% evaluation lies in figuring out what sort of query workload to use ---
% the performance analysis tools we have looked at suggest using an
% input file with somewhere on the order of one million queries, which
% is more than can feasibly be generated by hand. However, there does
% not appear to be a standard benchmark for this testing as there are
% for many other systems. We will likely generate large test files in a
% randomized way, but do not have a good sense of how to make these
% files accurately reflect realistic usage.

@@

\begin{lstlisting}
domain_name parse_domain_name(cursor cur, int[] data)
//@requires cur != NULL;
//@requires cursor_get(cur) < \length(data) * 32;
//@requires cursor_get(cur) % 8 == 0;
{
    int orig_ix = cursor_get(cur);
    int len = read_bits(cur, data, 8);

    if (len == 0) {
        return domain_nil();
    }

    if (len >= 0xc0) {
        // case: pointer
        // read the last 14 bits of this byte as a
        // pointer into the message (in octets)
        cursor_set(cur, orig_ix + 2);
        int ptr = read_bits(cur, data, 14);

        cursor ptrCursor = cursor_new(ptr * 8);
        return parse_domain_name(ptrCursor, data);
    }

    // now we parse an ordinary label; we retreat by one octet
    // so that read_string can see the length cell
    cursor_retreat(cur, 8);
    string lbl = read_string(cur, data);
    return domain_cons(lbl, parse_domain_name(cur, data));
}
\end{lstlisting}
